\documentclass[letterpaper,10pt]{article}

\setlength{\parindent}{0in}
\setlength{\parskip}{2ex}

%opening
\title{TTCS Assignment 2}
\author{Truman Ellis}

\begin{document}

\maketitle

\section*{Exercise 1}
The entries in /usr/share/dict/words are formatted one dictionary word per line
in the order A-Z, a-z, and special characters.

\section*{Exercise 2}
\subsection*{Part a}
\begin{verbatim}
for LETTER in {a..z}; do echo -n $LETTER:; grep -i "^$LETTER" words
| wc -l; done

a:5902
b:6095
c:9507
d:5811
e:3813
f:4075
g:3452
h:3856
i:3613
j:1259
k:1247
l:3437
m:5952
n:2015
o:2240
p:7539
q:464
r:5285
s:11072
t:5002
u:1899
v:1584
w:2711
x:56
y:380
z:287
\end{verbatim}

\subsection*{Part b}
\verb=grep -i "s$" words | wc -l=

46104 words end in \verb=s=

\verb=grep -i "'s$" words | wc -l=

24431 words end in \verb='s=

\subsection*{Part c}
\verb=grep -i "foo[^tdl]" words | wc -l=

7 words contain \emph{foo}, but not \emph{food, fool,} or \emph{foot}.

\subsection*{Part d}
Find the top 7 longest words in \verb=words=.

\verb=cat words | awk '{print length, $0}' | sort -nr | head -7=
\begin{verbatim}
23 electroencephalograph's
22 electroencephalographs
22 electroencephalogram's
22 counterrevolutionary's
22 counterrevolutionaries
22 Andrianampoinimerina's
21 electroencephalograph
\end{verbatim}

\section*{Exercise 3}
\begin{description}
\item[/h1 and /h2] Home directory, for all code developement, reports, and
personal data. Network filesystem, acked up daily.
\item[/org] Appears to be where specific applications go that are used by
various groups and centers. Network filesystem, backed up daily.
\item[/opt] Provides applications / compilers. Read only.
\item[/work] For generating shared data files. Network filesystem, not backed
up.
\item[/workspace] For generating data files on the local machine, much faster.
Local filesystem, not backed up.
\end{description}

\section*{Exercise 4}
It took 0.093 seconds to write 10MB and 0.556 seconds to write
100MB to \verb=/workspace=, while it takes 0.914 seconds to write 10MB and 9.094
seconds to write 100MB to \verb=/h2/truman=. It is faster to write to
\verb=/workspace= because it is a local disk, writing to my home directory has
to go over the network.

\section*{Exercise 5}
\subsection*{Part a}
\begin{verbatim}
du -hs

find . -type f | wc -l

find . -type f | grep '.c$' | wc -l
\end{verbatim}

The linux-2.6.24.1 directory is 304MB with 23062 files, of which 9748 are
``.c'' files.

\subsection*{Part b}
\begin{verbatim}
find . -type d | wc -l

ls -FR | grep '/$' | wc -l

tree -d
\end{verbatim}
According to \verb=ls= and \verb=tree=, there are 1355 directories. \verb=find=
counts 1356 because it also counts the current directory, ``\verb=.='' in it's
list.

\subsection*{Part c}
\begin{verbatim}
wc -l `find . -printf '%s %p\n' | sort -nr | head -1 | cut -f2 -d' '`

13947 ./fs/nls/nls_cp949.c
\end{verbatim}
The file \verb=fs/nls/nls_cp949.c= is the largest with 13947 newlines.

\subsection*{Part d}
\verb=wc -l `find . -type f | grep '.[ch]$'`=

counts 7803521 newlines in all \verb=.c= and \verb=.h= files.

\verb=awk 'NF {x++}END{print x}' `find . -type f | grep '.[ch]$'`=

counts 6737091 non-blank lines.

\subsection*{Part e}
How many lines in the Linux kernel make reference to ``happymeal''

\verb=egrep -ir "(happymeal)" * | wc -l=

There are 152 lines that have the upper, or lower case string ``happymeal''

\section*{Exercise 6}
My answer from last time was:
\begin{verbatim}
getent passwd | cut -f5 -d: | awk 'BEGIN{x="wwwwwwwwwwwww"}/[[:alpha:]]+\
[[:alpha:]]+/ {if (length < x && length >1){x=length;name=$0}}END{print name}'
\end{verbatim}
The shortest name of an ICES user is ``Jun Li''

But now I understand the use of backticks and your solution which found all
the shortest user names:
\begin{verbatim}
getent passwd | cut -f5 -d: | awk '{ if (NF==2 && length=='`getent passwd | cut
-f5 -d: | awk '{if(NF==2) {print 1+length($1)+length($2)} }' | sort -n | head
-1`') {print $1, $2}}' 
\end{verbatim}
which returns
\begin{verbatim}
Jun Li
Na Sai
Na Lei
Eh Tan
Ju Liu
J Shim
\end{verbatim}


\end{document}
